---
title: "s2s data story"
author: "Isaac Stone Simonelli"
date: "2024-04-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library (readxl)
library(tidyverse)
library(dplyr)
library(janitor)
library(ggplot2)
```

## R Markdown

We're doing a bit of a silly story based on the data for the ski to sea registration event. First step is to read that information in from the excel sheet provided.


```{r}

s2s <- read.csv("All Racers May 8.csv")

```

This is a very clean spreadsheet that we're dealing with, but let's go ahead tidy it up.


```{r}

s2s <- s2s %>%
  clean_names()

```

## Swag variables

We're going to count how many sizes of each sized shirt were requested. That's in the "T-Shirt Size" column. But first we need to select only people who got shirts. So first step is to take a look at the swag column, where we have people who got bennies, nothing or a t-shirt. Because we're not we're not distiquishing between shirt types we need to do it this way.


```{r}

value_counts <- s2s %>%
  mutate(category = case_when(
    merchandise_item == "Beanie Hat" ~ "Beanie Hat",
    merchandise_item == "Event T-Shirt" ~ "Shirt",
    merchandise_item == "No merch for me, please donate funds to Bellingham Youth Paddle Sports" ~ "Donate to Paddle Sports",
    is.na(merchandise_item) ~ "NA",
    TRUE ~ "Other"
  )) %>%
  count(category)

```

Let's get this broken down by percentage

```{r}

value_counts <- s2s %>%
  mutate(category = case_when(
    merchandise_item == "Beanie Hat" ~ "Beanie Hat",
    merchandise_item == "Event T-Shirt" ~ "Shirt",
    merchandise_item == "No merch for me, please donate funds to Bellingham Youth Paddle Sports" ~ "Donate to Paddle Sports",
    is.na(merchandise_item) ~ "NA",
    TRUE ~ "Other"
  )) %>%
  count(category) %>%
  mutate(percentage = n / sum(n) * 100) %>%
  print()


```


So that's kind of interesting and fun. So the N/A represents people who have yet to fill out the registration form but their team has signed up. We'll want to explain that in the story.

Now let's filter for sizes of shirts:

```{r}

shirt_db <- s2s %>%
  filter(merchandise_item == "Event T-Shirt")

```

We're going to do another counting exercise but this time based on shirt sizes

```{r}

 size_counts <- shirt_db %>%
  group_by(t_shirt_size) %>%
  count()

```

Let's see if we can get a breakdown based on percentages to compare.

Though it varies slightly depending on what t-shirt vendor is supplying the information, the average distribution suggested is 2% xs, 20% small, 28% medium, 27% large, 16% XL, 5% 2XL, and 3Xl

In total, it breaks down like this:
XS: 1 percent.
S: 7 percent.
M: 28 percent.
L: 30 percent.
XL: 20 percent.
2XL: 12 percent.
3XL: 2 percent.

```{r}

size_percentage <- shirt_db %>%
  group_by(t_shirt_size) %>%
  summarise(count = n()) %>%
  mutate(percentage = count / sum(count) * 100) %>%
  print()

```

## Demographic Data

We also need some of the more generic demographic data broken down for the story?

```{r}

gender_counts <- s2s %>%
  mutate(category = case_when(
    gender == "male" ~ "male",
    gender == "female" ~ "female",
    gender == "non_binary" ~ "non_binary",
    is.na(merchandise_item) ~ "NA",
    TRUE ~ "N/A"
  )) %>%
  count(category)

```

Let's get this broken down by percentage

```{r}

value_counts <- s2s %>%
  mutate(category = case_when(
    gender == "male" ~ "male",
    gender == "female" ~ "female",
    gender == "non_binary" ~ "non_binary",
    is.na(merchandise_item) ~ "NA",
    TRUE ~ "N/A"
  )) %>%
  count(category) %>%
  mutate(percentage = n / sum(n) * 100) %>%
  print()


```


Now let's get a breakdown of age groups:

```{r}

# Count the number of each value in the "age" column
age_counts <- s2s %>%
  count(age, name = "count") %>%
  arrange(desc(count)) %>%
  slice_head(n = 5)

# Print the result
print(age_counts)

```

Let's also grab the average age of participants:

```{r}

average_age <- s2s %>%
  summarise(mean_age = mean(age, na.rm = TRUE))

print(average_age)

```

We'll also go ahead and figure out what percent of folks are from Bellingham. Got to play into that hometown vibe.

```{r}

# Calculate the percentage of "Bellingham" values in the "city" column
bellingham_percentage <- s2s %>%
  summarise(percent_bellingham = mean(city == "Bellingham") * 100)

# Print the result
print(bellingham_percentage)


```

We're also going to want the total count:

```{r}

# Calculate the total number of "Bellingham" values in the "city" column
bellingham_count <- s2s %>%
  summarise(total_bellingham = sum(city == "Bellingham"))

# Print the result
print(bellingham_count)

```

For fun, let's see what we have from Seattle

```{r}

# Calculate the percentage of "Bellingham" values in the "city" column
seattle_percentage <- s2s %>%
  summarise(percent_seattle = mean(city == "Seattle") * 100)

# Print the result
print(seattle_percentage)

```

```{r}

seattle_count <- s2s %>%
  summarise(total_seattle = sum(city == "Seattle"))

# Print the result
print(seattle_count)

```


Let's take a look at how many people are coming in from out of county.

```{r}

out_of_country_percentage <- s2s %>%
  filter(country != "USA") %>%
  group_by(country) %>%
  summarise(percent = n() / nrow(s2s) * 100)

# Print the result
print(out_of_country_percentage)

```

## Published Story
The live version of the story can be found here: https://www.cascadiadaily.com/2024/may/23/ski-to-sea-by-the-numbers-3/












